---
title: "`ftw`: Plotting the data"
author: "Lorena Abad"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Plotting the data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Finally the `plot` method for the classes `ftw_atmp` and `ftw_tmp` are simpler than the explore method, given that the aim of the package is to provide an interactive way of exploring the data.

## Plot an object of class `ftw_atmp`

The `plot` method for this class, allows the user to have a plot of the network, and also analyze the trips according to the day of the week. Here are some examples:

```{r read atmp example, eval=TRUE, echo=TRUE, fig.width=6, fig.height=6}
library(ftw)
path <- system.file("extdata", package = "ftw")
data_atmp <- read_ftw(path, year = 2016, study_area = "Texel")
plot(data_atmp, type = "network")
plot(data_atmp, type = "trips per day")
```

## Plot an object of class `ftw_tmp`

At last, The `plot` method for this class is quite simple, but allows us to have a quick overview of the change between the year 2015 and 2016 for the network created by the users trips. 

```{r read tmp example, eval=TRUE, echo=TRUE, fig.width=6, fig.height=6}
library(ftw)
path <- system.file("extdata", package = "ftw")
data_tmp <- read_ftw(path, year = c(2015, 2016), study_area = "Smallingerland")
plot(data_tmp)
```
